The one characterizing the parental care stages

### 2019-10-11 analysis/07_hormones.Rmd
better timeline, thicker boxplots

sexcolors <- c("female" = "#F8766D", "male" = "#00BFC4")
charcolors <- c("control" = "#F8766D", "bldg" = "#D39200",
                "lay" =  "#93AA00", "inc_d3" = "#00BA38" ,"inc_d9" = "#00C19F" , "inc_d17" =  "#00B9E3", 
                "hatch" = "#619Cff" , "n5" =  "#DB72Fb", "n9" =  "#FF61C3")

@@

# rename some of the treatments levels, store in new columns
prolactin <- prolactin %>%
    mutate(sex = fct_recode(Sex,
                            "female" = "f",
                            "male" = "m"),
           treatment = fct_recode(Treatment,
                            "hatch" = "Hatch",
                            "inc_d17" = "Inc_d17",
                            "inc_d3" = "Inc_d3",
                            "inc_d9" = "Inc_d9",
                            "M_inc9" = "M_Inc9",
                            "M_inc3" = "M_Inc3",
                            "early" = "M_Inc8",
                            "early" = "M_inc8",
                            "M_inc17" = "M_Inc17",
                            "M_n2" = "M_hatch",
                            "control" = "baseline",
                            "n5" = "N5", 
                            "n9" = "N9"),
           study = fct_collapse(treatment,
                                 characterization = charlevels,
                                 manipulation = maniplevels,
                                removal = removelevels))

colnames(prolactin)[colnames(prolactin)=="Prolactin ng/mL"] <- "plasma_conc"
prolactin$hormone <- "prolactin"
prolactin <- prolactin %>% select(study, treatment, sex, hormone, plasma_conc)   %>% drop_na()

@@

PETC <- PETC %>%
    mutate(sex = fct_recode(sex,
                            "female" = "f",
                            "male" = "m"),
           study = fct_collapse(stage,
                                characterization = charlevels,
                                manipulation = maniplevels)) %>% 
  drop_na()
colnames(PETC)[colnames(PETC)=="stage"] <- "treatment"
PETC <- PETC %>% select(study, treatment, sex, hormone, plasma_conc)
summary(PETC)
```

```{r combine-clean-split}
hormones <- rbind(prolactin, PETC)
hormones$treatment <- factor(hormones$treatment, levels = combolevels)

hormones <- hormones %>%  mutate(treatment.sex = paste(treatment, sex, sep = " ")) 
hormones$treatment.sex <- factor(hormones$treatment.sex, levels = charsexlevels)

@@

prl.char <- hormones %>% filter(hormone == "prolactin", treatment %in% charlevels)   %>%  droplevels()
test.char <- hormones %>% filter(hormone == "testosterone", treatment %in% charlevels)   %>%  droplevels()
est.char <- hormones %>% filter(hormone == "estradiol", treatment %in% charlevels)   %>%  droplevels()
prog.char <- hormones %>% filter(hormone == "progesterone", treatment %in% charlevels)   %>%  droplevels()
cort.char <- hormones %>% filter(hormone == "cort", treatment %in% charlevels)   %>%  droplevels()

@@

  hormones %>% 
    filter(study == "characterization",
           hormone %in% c(myhormone))  %>% 
  ggplot(aes(x = treatment, y = plasma_conc, fill = treatment, color = sex)) +
    geom_boxplot(aes(outlier.colour = treatment)) +
    theme(axis.text.x =  element_blank(),
          legend.position = "none") +
    scale_fill_manual(values = mycolors) +
    scale_color_manual(values = sexcolors) +
    labs(y = myylab, x = NULL) 
}

@@

plot_grid(get_legend(a + theme(legend.position = "bottom")))
```

```{r aov-char}
aovSexTretment <- function(mydata, whichormone){
  aov2 <- aov(data = mydata, plasma_conc ~ treatment + sex)
  print(whichormone)
  print(summary(aov2))
  #print(TukeyHSD(aov2, which = "treatment"))
}

aovTretment  <- function(mydata, whichormone){
  aov1 <- aov(data = mydata, plasma_conc ~ treatment )
  print(whichormone)
  print(summary(aov1))
  #print(TukeyHSD(aov1, which = "treatment"))
}

@@

  hormones %>% 
    filter(treatment %in% manipchar,
           hormone %in% c(myhormone))  %>% 
  ggplot(aes(x = treatment, y = plasma_conc, fill = sex)) +
    geom_boxplot() + 
    theme(axis.text.x = element_text(angle = 45, hjust = 1),
          legend.position = "none",
          strip.text = element_blank()) +
    scale_fill_manual(values = mycolors) +
    labs(y = myylab, x = NULL) +
    facet_wrap(~sex) +
    annotate("rect", xmin = 0.5, xmax = 1.5, ymin = 0, ymax = 125, alpha = .2) +
    annotate("rect", xmin = 3.5, xmax = 5.5, ymin = 0, ymax = 125, alpha = .2) 
}

@@

  hormones %>% 
    filter(treatment %in% manipchar,
           hormone %in% c(myhormone))  %>% 
  ggplot(aes(x = treatment, y = plasma_conc, fill = sex)) +
    geom_boxplot() + 
    theme(axis.text.x = element_text(angle = 45, hjust = 1),
          legend.position = "none",
          strip.text = element_blank()) +
    scale_fill_manual(values = mycolors) +
    labs(y = myylab, x = NULL) +
    facet_wrap(~sex) +
    annotate("rect", xmin = 2.5, xmax = 4.5, ymin = 0, ymax = myymax, alpha = .2)
}

@@

  hormones %>% 
    filter(treatment %in% removechar,
           hormone %in% c(myhormone))  %>% 
  ggplot(aes(x = treatment, y = plasma_conc, fill = sex)) +
    geom_boxplot() + 
    theme(axis.text.x = element_text(angle = 45, hjust = 1),
          legend.position = "none",
          strip.text = element_blank()) +
    scale_fill_manual(values = mycolors) +
    labs(y = myylab, x = NULL) +
    facet_wrap(~sex) +
    annotate("rect", xmin = 1.5, xmax = 2.5, ymin = 0, ymax = 125, alpha = .2) +
    annotate("rect", xmin = 3.5, xmax = 4.5, ymin = 0, ymax = 125, alpha = .2) +
    annotate("rect", xmin = 5.5, xmax = 6.5, ymin = 0, ymax = 125, alpha = .2) + 
    annotate("rect", xmin = 7.5, xmax = 8.5, ymin = 0, ymax = 125, alpha = .2) 
}

>>>>

sexcolors <- c("female" = "#F8766D", "male" = "#00BFC4")
charcolors <- c("control" = "#F8766D", "bldg" = "#D39200",
                "lay" =  "#93AA00", "inc_d3" = "#00BA38" ,
                "inc_d9" = "#00C19F" , "inc_d17" =  "#00B9E3", 
                "hatch" = "#619Cff" , "n5" =  "#DB72Fb", "n9" =  "#FF61C3")

@@

# rename some of the timepoints levels, store in new columns
prolactin <- prolactin %>%
    mutate(sex = fct_recode(Sex,
                            "female" = "f",
                            "male" = "m"),
           timepoint = fct_recode(Treatment,
                            "hatch" = "Hatch",
                            "inc_d17" = "Inc_d17",
                            "inc_d3" = "Inc_d3",
                            "inc_d9" = "Inc_d9",
                            "M_inc9" = "M_Inc9",
                            "M_inc3" = "M_Inc3",
                            "early" = "M_Inc8",
                            "early" = "M_inc8",
                            "M_inc17" = "M_Inc17",
                            "M_n2" = "M_hatch",
                            "control" = "baseline",
                            "n5" = "N5", 
                            "n9" = "N9"),
           study = fct_collapse(timepoint,
                                 characterization = charlevels,
                                 manipulation = maniplevels,
                                removal = removelevels))

colnames(prolactin)[colnames(prolactin)=="Prolactin ng/mL"] <- "plasma_conc"
prolactin$hormone <- "prolactin"
prolactin <- prolactin %>% select(study, timepoint, sex, hormone, plasma_conc)   %>% drop_na()

@@

PETC <- PETC %>%
    mutate(sex = fct_recode(sex,
                            "female" = "f",
                            "male" = "m"),
           study = fct_collapse(stage,
                                characterization = charlevels,
                                manipulation = maniplevels)) %>% 
  drop_na()
colnames(PETC)[colnames(PETC)=="stage"] <- "timepoint"
PETC <- PETC %>% select(study, timepoint, sex, hormone, plasma_conc)
summary(PETC)
```

```{r combine-clean-split}
hormones <- rbind(prolactin, PETC)
hormones$timepoint <- factor(hormones$timepoint, levels = combolevels)

hormones <- hormones %>%  mutate(timepoint.sex = paste(timepoint, sex, sep = " ")) 
hormones$timepoint.sex <- factor(hormones$timepoint.sex, levels = charsexlevels)

@@

prl.char <- hormones %>% filter(hormone == "prolactin", timepoint %in% charlevels)   %>%  droplevels()
test.char <- hormones %>% filter(hormone == "testosterone", timepoint %in% charlevels)   %>%  droplevels()
est.char <- hormones %>% filter(hormone == "estradiol", timepoint %in% charlevels)   %>%  droplevels()
prog.char <- hormones %>% filter(hormone == "progesterone", timepoint %in% charlevels)   %>%  droplevels()
cort.char <- hormones %>% filter(hormone == "cort", timepoint %in% charlevels)   %>%  droplevels()

@@

  hormones %>% 
    filter(study == "characterization",
           hormone %in% c(myhormone))  %>% 
  ggplot(aes(x = timepoint, y = plasma_conc, fill = timepoint, color = sex)) +
    geom_boxplot(aes(outlier.colour = timepoint), lwd=1.05) +
    theme(axis.text.x =  element_blank(),
          legend.position = "none") +
    scale_fill_manual(values = mycolors) +
    scale_color_manual(values = sexcolors) +
    labs(y = myylab, x = NULL) +
    guides(fill = guide_legend(order=1),
         color = guide_legend(order=2))
}

@@

```{r aov-char}
aovSexTretment <- function(mydata, whichormone){
  aov2 <- aov(data = mydata, plasma_conc ~ timepoint + sex)
  print(whichormone)
  print(summary(aov2))
  #print(TukeyHSD(aov2, which = "timepoint"))
}

aovTretment  <- function(mydata, whichormone){
  aov1 <- aov(data = mydata, plasma_conc ~ timepoint )
  print(whichormone)
  print(summary(aov1))
  #print(TukeyHSD(aov1, which = "timepoint"))
}

@@

  hormones %>% 
    filter(timepoint %in% manipchar,
           hormone %in% c(myhormone))  %>% 
  ggplot(aes(x = timepoint, y = plasma_conc, fill = sex)) +
    geom_boxplot() + 
    theme(axis.text.x = element_text(angle = 45, hjust = 1),
          legend.position = "none",
          strip.text = element_blank()) +
    scale_fill_manual(values = mycolors) +
    labs(y = myylab, x = NULL) +
    facet_wrap(~sex) +
    annotate("rect", xmin = 0.5, xmax = 1.5, ymin = 0, ymax = 125, alpha = .2) +
    annotate("rect", xmin = 3.5, xmax = 5.5, ymin = 0, ymax = 125, alpha = .2) 
}

@@

  hormones %>% 
    filter(timepoint %in% manipchar,
           hormone %in% c(myhormone))  %>% 
  ggplot(aes(x = timepoint, y = plasma_conc, fill = sex)) +
    geom_boxplot() + 
    theme(axis.text.x = element_text(angle = 45, hjust = 1),
          legend.position = "none",
          strip.text = element_blank()) +
    scale_fill_manual(values = mycolors) +
    labs(y = myylab, x = NULL) +
    facet_wrap(~sex) +
    annotate("rect", xmin = 2.5, xmax = 4.5, ymin = 0, ymax = myymax, alpha = .2)
}

@@

  hormones %>% 
    filter(timepoint %in% removechar,
           hormone %in% c(myhormone))  %>% 
  ggplot(aes(x = timepoint, y = plasma_conc, fill = sex)) +
    geom_boxplot() + 
    theme(axis.text.x = element_text(angle = 45, hjust = 1),
          legend.position = "none",
          strip.text = element_blank()) +
    scale_fill_manual(values = mycolors) +
    labs(y = myylab, x = NULL) +
    facet_wrap(~sex) +
    annotate("rect", xmin = 1.5, xmax = 2.5, ymin = 0, ymax = 125, alpha = .2) +
    annotate("rect", xmin = 3.5, xmax = 4.5, ymin = 0, ymax = 125, alpha = .2) +
    annotate("rect", xmin = 5.5, xmax = 6.5, ymin = 0, ymax = 125, alpha = .2) + 
    annotate("rect", xmin = 7.5, xmax = 8.5, ymin = 0, ymax = 125, alpha = .2) 
}

### 2019-10-11 docs/ms1-charactization.Rmd
better timeline, thicker boxplots

## Fig 2: hormones

@@

## Figure 3: general rnaseq pattern, tissue > sex > stage

_Needs to be replaced with a better PCA plot_

@@

@@

## Figure 5: PRL in prolatin mirrors PRL in bloo

>>>>

## Fig 2: prolactin flucuates the most and peaks near hatch in males and females

@@

## Figure 3: general rnaseq pattern, tissue > sex > parental stage

_Needs to be replaced with a better PCA plot showing all samples as points_

@@

_redo color scheme, fix titiles_

@@

## Figure 5: PRL gene expression in the pituitary in prolatin mirrors PRL in blood
